
\newglossaryentry{linear system}
{type=\thisgls,
name={linear system},
description={A linear system is a system which is linear.}
symbol={}
}

\newglossaryentry{first-order model}
{
type=\thisgls,
name={first-order model},
description={A first-order, linear, time-invariant ordinary differential equation.}
}
\newglossaryentry{time constant}
{
type=\thisgls,
name={first-order model},
description={For a first-order model with a step input, the time duration from the initiation of the step to the time when the model output is 63.2\% of the way to the steady-state value.},
symbol={\ensuremath{\tau}}
}

\newglossaryentry{mathematical model}
{
type=\thisgls,
name={first-order model},
description={An equation or set of equations that behave in a similar way to a physical system under certain conditions.}
}

\newglossaryentry{step function}
{
type=\thisgls,
name={step input},
description={A mathematical function that is zero for all time less than $t=0$ and unity for all time greater than or equal to $t=1$.},
symbol={\ensuremath{\mu(t)}}
}

\newglossaryentry{step response}
{
type=\thisgls,
name={step response},
description={The output of a mathematical function, as a function of time, when a step input is given as the input or forcing function.}
}

\newglossaryentry{steady-state response}
{
type=\thisgls,
name={steady-state response},
description={The output of a mathematical model or physical system when in equilibrium.  The steady-state response is often contrasted with the transient response.}
}
\newglossaryentry{superposition}
{
type=\thisgls,
name={superposition},
description={The property of linear mathematical models that states that the total response of the system caused by two (or more) inputs is the sum of the responses to each input considered independently}
}

\newglossaryentry{second-order model}
{
type=\thisgls,
name={second-order model},
description={A linear, ordinary, time-invariant, second-order differential equation.  For our purposes we assume the model is underdamped or undamped.}
}

\newglossaryentry{damping ratio}
{
type=\thisgls,
name={damping ratio},
description={``The damping ratio is a dimensionless measure describing how oscillations in a system decay after a disturbance''---from Wikipedia!},
symbol={$\zeta$}
}

\newglossaryentry{undamped natural frequency}
{
type=\thisgls,
name={undamped natural frequency},
description={The theoretical frequency of oscillation (in \unitfrac[]{rad}{s}) of the response of a second-order model if there was no damping (loss) in the system.},
symbol={$\omega_n$}
}

\newglossaryentry{damped natural frequency}
{
type=\thisgls,
name={damped natural frequency},
description={The actual frequency of oscillation (in \unitfrac[]{rad}{s} of the response of a second-order model with damping (loss) included: $\omega_d = \omega_n \sqrt{1-\zeta^2}$},
symbol={$\omega_d$}
}
